---
title: "Using the ABCT Analysis Pipeline"
output: 
  rmarkdown::html_vignette: 
    toc: true
    self_contained: true
fig_width: 10 
fig_height: 6 
vignette: >
  %\VignetteIndexEntry{Using the ABCT Analysis Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style>
p.caption {
  font-size: 1.5em;
}
</style>
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

This vignette demonstrates how to identify malignant cells and perform anchor-based cell type classification using the CosMx SMI NSCLC FFPE Dataset, Lung 5-1 sample. We employ the FindMalignantCells function to detect malignant cells based on predefined marker genes and subsequently apply RunABCT for cell type annotation using an anchor-based classification approach. This workflow showcases the integration of spatial transcriptomics data with computational methods for accurate cell type identification.

### Loading Required Libraries

Ensure that all necessary libraries are installed and loaded.

```{r libraries}
library(Seurat)
library(SeuratWrappers)
library(Banksy)
library(harmony)
library(dplyr)
library(scales)
library(EnvStats)
library(stringr)
library(matrixStats)
library(InSituType)
library(UCell)
library(ggplot2)
library(MASS)
library(pracma)

source("./ABCT.r")
```

### Loading and Preparing Data

Load the data for analysis. Here, we demonstrate using a Seurat object. The processed data can be downloaded from https://kbds.re.kr/hissta/datasetinfo?sampleIdx=19.

```{r load-data}
obj <- readRDS(paste0("/path/to/your/data/", "obj_final.rds"))
```

### Define Cell Type List and Marker Genes

We define a list of cell types and specify marker genes for the analysis.

```{r cell-types}
celltype_list <- c("Malignant", "Epithelial", "CD4T", "CD8T", "NK", "B", "Plasma",
                   "Macrophage", "Monocyte", "DC", "Mast", "Neutrophil", 
                   "Endothelial", "Fibroblast", "Unknown")

names(celltype_list) <- c(
  "#bf4040", "#FB8072", "#8DD3C7", "#FFFFB3", "#BEBADA", "#3288bd", 
  "#FDB462", "#B3DE69", "#FCCDE5", "#84cdee", "#BC80BD", "#CCEBC5", 
  "#FFED6F", "#1b9e77", "gray"
)

malignant_marker_list <- data.frame(
  cluster = "Malignant",
  gene = c("SOX9", "FGFR1", "KRAS", "MYC", "EGFR")
)

ABCT_marker_list <- read.csv("/path/to/marker_list.csv")
ABCT_marker_list$cluster <- factor(ABCT_marker_list$cluster, levels = celltype_list)
ABCT_marker_list <- ABCT_marker_list %>% arrange(cluster)
```

### Identifying Malignant Cells

Use the `FindMalignantCells` function to identify malignant cells in the data.

```{r find-malignant}
malignant_path <- "/path/to/output/malignant/"

obj <- FindMalignantCells(
  obj,
  assay = "SCT",
  ctrl_assay = "negprobes",
  marker_list = malignant_marker_list,
  use_spatial = TRUE,
  M = 1,
  lambda = 0.2,
  area = NULL,
  w_neg = 0,
  dimx = "x_global_px",
  dimy = "y_global_px",
  smooth_reduction = "spatial_pca",
  path = malignant_path
)
```

### Performing ABCT Classification

Subsequently, classify the non-malignant cells using the RunABCT function. By default, if a BANKSY object already exists in the Seurat object, RunABCT will use the existing BANKSY object rather than running it again. If you wish to rerun BANKSY on a subsetted object, you should remove the existing BANKSY assay before executing the function.

```{r run-abct}
subobj <- subset(obj, subset = malignant_result != "Malignant")

abct_path <- "/path/to/output/abct/"

subobj <- RunABCT(
  subobj,
  assay = "SCT",
  ctrl_assay = "negprobes",
  marker_list = ABCT_marker_list,
  color_list = names(celltype_list[celltype_list %in% ABCT_marker_list$cluster]),
  method = "quantile",
  use_spatial = TRUE,
  M = 1,
  lambda = 0.2,
  dimx = "x_global_px",
  dimy = "y_global_px",
  smooth_reduction = "spatial_pca",
  path = abct_path
)
```

### Updating Metadata

Finally, update the metadata in the original object with the ABCT classification results.

```{r update-metadata}
obj <- update_metadata(obj, subobj, celltype_list)
```

<br>
<br>

<details>
<summary><strong>Session Info</strong></summary>
```{r sessioninfo, echo=FALSE}
sessionInfo()
```
</details>

<br>
<br>
